### 2018-03-31 cav2018/rebuttal.txt

Review1

Q1: ”Were you able to find any bugs in existing solvers with your tool? [...] Is the tool more suitable for performance analysis, rather than correctness testing?”

A1: StringFuzz can be used for both, but is more suitable for performance analysis. We used it as such when working with Z3str3, Z3str2 and CVC4. For example it helped find actual implementation bugs (e.g. commits 6308636 to 3865c45, and 66bc68f to 7b536e9) in previous versions of Z3str3, and also helped find performance issues in Z3str3 and CVC4.

Q2: “I am not sure how useful the modified industrial sets are... Is it intended for correctness testing or performance tuning? Solvers are expected to take advantage of structural properties [...] Destroying such properties might lead the solvers astray, but I am not sure if one can learn anything from such experiments...”

A2: Modified industrial instances were useful for performance analysis. Fuzzing them with StringFuzz allowed us to maintain instance structure in some respects, but also make the instance extremal in others. For example in Amazon instances we replaced literals with unprintable characters, and scaled the lengths of literals, without changing any other aspect of the instance.

Review2

Q1: “[...] I would expect some form of evolutionary algorithms [...]”

A1: Thank you for this suggestion! In fact, we’re already working on augmenting StringFuzz with an evolutionary approach.

Q2: “Further, the description of some of the generators and transformers is very brief [...]”

A2: This was intentional, to save space. We provide in-depth generator descriptions/examples on the StringFuzz website[1].

Q3: “[...] both the entire framework as well as the generators and transforms seem to be still relatively simple.”

A3: We respectfully disagree with this point. Some of the transformations are satisfiability-preserving and are fairly complex.

Q4: “Page 3, [...] (a) there should be no indentation there, (b) RegEx is not defined.”

A4: We will fix the indentation and define RegEx.

Q5: “Page 4, [...] (a) two cases of wrong indentation, (b) not clear enough: in particular, can concatenation appear lower in the terms too?”

A5: Indeed, concatenation can only appear on the top level. We will clarify this, and also fix the indentation.

Review3

Q1: “It would be nice to consider extending the fuzzing to cover other theories as well, such as bit-vectors.”

A1: We agree, and are currently extending the fuzzer to support the entire SMT-LIB 2.6 standard.

Review4

Q1: ”Section 3 describes a suite of over 10000 string benchmarks. Section 4 promises an experimental evaluation, but presents results on only a tiny fraction of these benchmarks, including none of the industrial benchmarks. There is no explanation as to why so few results are reported. I think that the paper is pretty weak without a full set of experimental results on the new benchmarks”

A1: We indeed performed experiments on all the suites. We presented the suites that showed interesting behaviour, and for brevity omitted the ones that showed all solvers performing similarly. We currently present all the data and graphs on our website[2]. We will add them to the Appendix.

Q2: “do the tools generally agree on an input syntax?”

A2: They generally agree, but we had to account for two versions of the string syntax. StringFuzz supports both using the --language-in and --language-out flags.

Q3: “The subsection on Regex generating capabilities is rather informal and quite unclear.  Are you mixing SMT-LIB syntax with regular expression syntax at the meta-level?  It's very confusing.”

A3: It is indeed confusing. We will rewrite the example in SMT-LIB syntax only.

Q4: “You mention two data sets, "Concats-Big" and "Concats-Extracts-Small", both of which are not listed in Table 2a.  Where do these data sets come from?”

A4: We grouped Concats-Big and Concats-Small under Concats (and same for Concats-Extracts) in the table for brevity, but that is indeed unclear. We will add them in the table explicitly.

Review5

Q1: “I ran this script [...] but did not see any errors after 30 minutes or so. It would have been nice to say if the StringFuzz tool was able to find any crashing inputs for the string solvers.”

A1: StringFuzz has many options (visible with the --help flag) that may speed up the time to crash. For example “stringfuzzg --random random-ast --meaningful --num-vars 2 --num-asserts 10 --depth 3“ will yield more challenging instances. We discovered errors in Z3str3 with similar scripts, but have not fully explored other solvers.

Review6

Q1: “They didn't manage to build the solver Z3str2 on the VM.”

A1: Indeed, we failed to do it in time for the submission. We apologise.

Q2: “I was surprised to see the checksum for the machine image.”

A2: Indeed, the checksum is invalid when it’s on the VM. It was transferred by mistake. We apologise for the oversight.

[1] http://stringfuzz.dmitryblotsky.com/problems/
[2] http://stringfuzz.dmitryblotsky.com/results/latest/

### 2018-04-07 includes/introduction.tex
Need to push contributions to second page, but other than that it is getting there

\subsubsection{Contributions}

\begin{enumerate}
    \item \textbf{The \fuzzer{} tool}:
        In Sect.~\ref{sec:fuzzer}, we describe a modular fuzzer that can
        transform and generate \smtfull{} string and regex
        instances.\footnote{We assume basic
        familiarity with string solvers and their input
        language.} Scaling inputs (e.g., long string constants,
        deep concatenations) are particularly useful in identifying asymptotic
        behaviors in solvers, and \fuzzer{} has many options to generate them.
        We briefly document \fuzzer{}'s components and modular architecture.
        We provide example use cases to demonstrate its utility as an
        exploratory solver testing tool.

>>>>

\subsubsection{Contributions}\footnote{All source code, problem suites, and supplementary material
referenced in this paper are available at the \fuzzer{} website \cite{website}.}
\begin{enumerate}
    \item \textbf{The \fuzzer{} tool}:
        In Sect.~\ref{sec:fuzzer}, we describe a modular fuzzer that can
        transform and generate \smtfull{} string and regex
        instances. Scaling inputs (e.g., long string constants,
        deep concatenations) are particularly useful in identifying asymptotic
        behaviors in solvers, and \fuzzer{} has many options to generate them.
        We briefly document \fuzzer{}'s components and modular architecture.
        We provide example use cases to demonstrate its utility as an
        exploratory solver testing tool.

### 2018-04-07 includes/stringfuzz.tex
Need to push contributions to second page, but other than that it is getting there

\fuzzer{} is implemented as a Python package, and comes with several
executables to generate, transform, and analyze \smtfull{} string and regex
instances. Its components are implemented as \unix{} ``filters'' to enable easy
integration with other tools (including themselves). For example, the
outputs of generators can be piped into transformers, and transformers
can be chained to produce a stream of tuned inputs to a
solver. \fuzzer{} is composed of the following tools:
\begin{description}
    \item[\generator{}] \hfill \\
    This tool generates \smt{} instances. It supports several generators and
    options that specify its output. Details can be found in
    Table~\ref{tbl:generators}.
    \item[\transformer{}] \hfill \\
    This tool transforms \smt{}
    instances. It supports several transformers and options that specify
    its output and input, which are explained in
    Table~\ref{tbl:transformers}. Note that transformers
    \textit{Translate} and \textit{Reverse} also preserve
    satisfiability under certain conditions.
    \item[\texttt{stringstats}] \hfill \\
    This tool takes an \smt{}
    instance as input and outputs its properties: the number of
    variables/literals, the max/median syntactic depth of expressions, the
    max/median literal length, etc.
\end{description}
We organized \fuzzer{} to be easily extended. As evidence of our success we note
that while the whole project
contains \linesInFuzzer{} lines of code, it takes an average of
\linesPerX{} lines of code to create a transformer. \fuzzer{} can either be
installed from source, or from the Python PIP package
repository.\footnote{The link to the source code will be added after double-blind review.}

\begin{table}[t]
    \caption{\fuzzer{} built-in (a) generators and (b) transformers.}
    \begin{subtable}{1\textwidth}
        \centering
        \caption{\generator{} built-in generators.}
        \label{tbl:generators}
        \begin{tabular}{ l l }
            \toprule
            \textbf{Name}
            & \textbf{Generates instances that have ...} \\
            \midrule
            \textit{Concats}
            & Long concats and optional random extracts. \\
            \textit{Lengths}
            & Many variables (and their concats) with length constraints. \\
            \textit{Overlaps}
            & An expression of the form A.X = X.B. \\
            \textit{Equality}
            & An equality among concats, each with variables or constants. \\
            \textit{Regex}
            & Regexes of varying complexity. \\
            \textit{Random-Text}
            & Random, likely syntactically \textit{in}valid text. \\
            \textit{Random-AST}
            & Random, but semantically \textit{valid} text. \\
            \bottomrule
        \end{tabular}
    \end{subtable}
    \begin{subtable}{1\textwidth}
        \centering
        \caption{\transformer{} built-in transformers.}
        \label{tbl:transformers}
        \begin{tabular}{l l}
            \toprule
            \textbf{Name}
            & \textbf{The transformer ...} \\
            \midrule
            \textit{Fuzz}
            & Replaces literals and operators with similar ones.\\
            \textit{Graft}
            & Randomly swaps non-leaf nodes with leaf nodes.\\
            \textit{Multiply}\footnote{Satisfiable inputs
            will produce satisfiable outputs (see Appendix).}
            & Multiplies integers and repeats strings by N.\\
            \textit{Nop}
            & Does nothing (can translate between \smtfull{}).\\
            \textit{Reverse}\footnote{Input and output
            instances will be equisatisfiable (see Appendix).}
            & Reverses all string literals and concat arguments.\\
            \textit{Rotate}
            & Rotates compatible nodes in syntax tree.\\
            \textit{Translate}\footnotemark[4]
            & Permutes the alphabet.\\
            \textit{Unprintable}
            & Replaces characters in literals with unprintable ones.\\
            \bottomrule
        \end{tabular}
    \end{subtable}
\end{table}

@@

where $R_i \in RegEx$, and $Min, Max \in Int$. More simply, the
instance is a set of one or more regex constraints on a single
variable, with optional maximum and minimum length constraints. The
regex constraints $R$ are each of the form:
\begin{align*}
    & \texttt{(re.++}\; T_0\; \texttt{(re.++}\; T_1\;
    \texttt{...}\; \texttt{(re.++}\; T_{n-1}\; T_n\; \texttt{))}
\end{align*}

and each $T_i$ is a recursive term of the form:
\begin{align*}
    & \texttt{(re.*}\; T_{i_j}\; \texttt{) | (re.+}\; T_{i_j}\;
    \texttt{) | (re.union}\; T_{i_{j_1}}\; T_{i_{j_2}}\; \texttt{)}
\end{align*}

where $j$ is the specified depth of recursion. Terms at depth 0 are
regex constants. Informally, this form describes a concatenation of
regex terms, where each term is a random nested regex operator (chosen from
regex Kleene star, repetition, and union), up to a specified depth,
terminating in a regex literal. Below are three example regex instances
(separated by spaces) of depth 2 produced by this scheme:
\begin{align*}
    & ((\texttt{a}|\texttt{b})|(\texttt{cc})+)\quad\quad
    ((\texttt{ddd})*)+\quad\quad ((\texttt{ee})+|(\texttt{fff})*)
\end{align*}

>>>>

\fuzzer{} is implemented as a Python package, and comes with several
executables to generate, transform, and analyze \smtfull{} string and regex
instances. Its components are implemented as \unix{} ``filters'' to enable easy
integration with other tools (including themselves). For example, the
outputs of generators can be piped into transformers, and transformers
can be chained to produce a stream of tuned inputs to a
solver. \fuzzer{} is composed of the following tools:
\begin{description}
    \item[\generator{}] \hfill \\
    This tool generates \smt{} instances. It supports several generators and
    options that specify its output. Details can be found in
    Table~\ref{tbl:generators}.
    \item[\transformer{}] \hfill \\
    This tool transforms \smt{}
    instances. It supports several transformers and options that specify
    its output and input, which are explained in
    Table~\ref{tbl:transformers}. Note that transformers
    \textit{Translate} and \textit{Reverse} also preserve
    satisfiability under certain conditions.
    \item[\texttt{stringstats}] \hfill \\
    This tool takes an \smt{}
    instance as input and outputs its properties: the number of
    variables/literals, the max/median syntactic depth of expressions, the
    max/median literal length, etc.
\end{description}
We organized \fuzzer{} to be easily extended. As evidence of our success we note
that while the whole project
contains \linesInFuzzer{} lines of code, it takes an average of
\linesPerX{} lines of code to create a transformer. \fuzzer{} can either be
installed from source, or from the Python PIP package repository.

\begin{table}[t]
    \caption{\fuzzer{} built-in (a) generators and (b) transformers.}
    \begin{subtable}{1\textwidth}
        \centering
        \caption{\generator{} built-in generators.}
        \label{tbl:generators}
        \begin{tabular}{ l l }
            \toprule
            \textbf{Name}
            & \textbf{Generates instances that have ...} \\
            \midrule
            \textit{Concats}
            & Long concats and optional random extracts. \\
            \textit{Lengths}
            & Many variables (and their concats) with length constraints. \\
            \textit{Overlaps}
            & An expression of the form A.X = X.B. \\
            \textit{Equality}
            & An equality among concats, each with variables or constants. \\
            \textit{Regex}
            & Regexes of varying complexity. \\
            \textit{Random-Text}
            & Random, likely syntactically \textit{in}valid text. \\
            \textit{Random-AST}
            & Random, but semantically \textit{valid} text. \\
            \bottomrule
        \end{tabular}
    \end{subtable}
    \begin{subtable}{1\textwidth}
        \centering
        \caption{\transformer{} built-in transformers.}
        \label{tbl:transformers}
        \begin{tabular}{l l}
            \toprule
            \textbf{Name}
            & \textbf{The transformer ...} \\
            \midrule
            \textit{Fuzz}
            & Replaces literals and operators with similar ones.\\
            \textit{Graft}
            & Randomly swaps non-leaf nodes with leaf nodes.\\
            \textit{Multiply}\footnote{Can guarantee satisfiable output instances 
            from satisfiable input instances \cite{website}.}
            & Multiplies integers and repeats strings by N.\\
            \textit{Nop}
            & Does nothing (can translate between \smtfull{}).\\
            \textit{Reverse}\footnote{Can guarantee input and output
            instances will be equisatisfiable \cite{website}.}
            & Reverses all string literals and concat arguments.\\
            \textit{Rotate}
            & Rotates compatible nodes in syntax tree.\\
            \textit{Translate}\footnotemark[6]
            & Permutes the alphabet.\\
            \textit{Unprintable}
            & Replaces characters in literals with unprintable ones.\\
            \bottomrule
        \end{tabular}
    \end{subtable}
\end{table}

@@

\noindent where $R_i \in RegEx$, and $Min, Max \in Int$. More simply, the
instance is a set of one or more regex constraints on a single
variable, with optional maximum and minimum length constraints. The
regex constraints $R$ are each of the form:
\begin{align*}
    & \texttt{(re.++}\; T_0\; \texttt{(re.++}\; T_1\;
    \texttt{...}\; \texttt{(re.++}\; T_{n-1}\; T_n\; \texttt{))}
\end{align*}

\noindent and each $T_i$ is a recursive term of the form:
\begin{align*}
    & \texttt{(re.*}\; T_{i_j}\; \texttt{) | (re.+}\; T_{i_j}\;
    \texttt{) | (re.union}\; T_{i_{j_1}}\; T_{i_{j_2}}\; \texttt{)}
\end{align*}

\noindent where $j$ is the specified depth of recursion. Terms at depth 0 are
regex constants. Informally, this form describes a concatenation of
regex terms, where each term is a random nested regex operator (chosen from
regex Kleene star, repetition, and union), up to a specified depth,
terminating in a regex literal. Below are three example regex instances
(separated by spaces) of depth 2 produced by this scheme:
\begin{align*}
    & ((\texttt{a}|\texttt{b})|(\texttt{cc})+)\quad\quad
    ((\texttt{ddd})*)+\quad\quad ((\texttt{ee})+|(\texttt{fff})*)
\end{align*}

### 2018-04-07 includes/suites.tex
fixed small mistake
Adjusting numbers.
Need to push contributions to second page, but other than that it is getting there

In this section, we describe the benchmark suites we generated
with \fuzzer{}, and on which we conducted our experimental
evaluation.\footnote{The link to the repository will be added after
double-blind review.} Table~\ref{tbl:generated} describes the string
and regex instances that were
generated by \generator{}. Table~\ref{tbl:transformed} lists benchmarks
derived from existing seed instances by iteratively applying \transformer{}.
Every transformed instance is named according to its
seed and the transformations it undertook. For example, input
\texttt{z3-regex-1-fuzz-graft.smt2} was transformed by applying
\textit{Fuzz} and then \textit{Graft} to \texttt{z3-regex-1.smt2}.

@@

\begin{table}[t]
    \caption{Repository of 15,843 \smtfull{} instances.}
    \begin{subtable}{1\textwidth}
        \centering
        \caption{\generator{}-generated instances.}
        \label{tbl:generated}
        \begin{tabular}{llr}
            \toprule
            \textbf{Name}
                & \textbf{Instances have a ...}
                & \textbf{Quantity}
            \\
            \midrule
            \textit{Concats}
                & Right-heavy, deep tree of concats.
                & 600 \\
            \textit{Concats-Balanced}
                & Balanced, deep tree of concats.
                & 100 \\
            \textit{Concats-Extracts}
                & Single concat tree, with character extractions.
                & 600 \\
            \textit{Lengths}
                & Single, large length constraint on a variable.
                & 1,000 \\
            \textit{Lengths-Concats}
                & Tree of fixed-length concats of variables.
                & 500 \\
            \textit{Overlaps}
                & Formula of the form A.X = X.B.
                & 100 \\
            \textit{Regex}
                & Complex regex membership test.
                & 1,200 \\
            \textit{Many-Regexes}
                & Multiple random regex membership tests.
                & 600 \\
            \textit{Regex-Deep}
                & Regex membership test with many nested operators.
                & 450 \\
            \textit{Regex-Pair}
                & Test for membership in one regex, but not another.
                & 600 \\
            \textit{Regex-Lengths}
                & Regex membership test, and a length constraint.
                & 600 \\
            \textit{Different-Prefix}
                & Equality of two deep concats with different prefixes.
                & 300 \\
            \bottomrule
        \end{tabular}
    \end{subtable}

>>>>

In this section, we describe the benchmark suites we generated
with \fuzzer{}, and on which we conducted our experimental
evaluation. Table~\ref{tbl:generated} describes the string
and regex instances that were
generated by \generator{}. Table~\ref{tbl:transformed} lists benchmarks
derived from existing seed instances by iteratively applying \transformer{}.
Every transformed instance is named according to its
seed and the transformations it undertook. For example,
\texttt{z3-regex-1-fuzz-graft.smt2} was transformed by applying
\textit{Fuzz} and then \textit{Graft} to \texttt{z3-regex-1.smt2}.

@@

\begin{table}[t]
    \caption{Repository of 10,258 \smtfull{} instances.}
    \begin{subtable}{1\textwidth}
        \centering
        \caption{\generator{}-generated instances.}
        \label{tbl:generated}
        \begin{tabular}{llr}
            \toprule
            \textbf{Name}
                & \textbf{Instances have a ...}
                & \textbf{Quantity}
            \\
            \midrule
            \textit{Concats}
                & Right-heavy, deep tree of concats.
                & 120 \\
            \textit{Concats-Balanced}
                & Balanced, deep tree of concats.
                & 100 \\
            \textit{Concats-Extracts}
                & Single concat tree, with character extractions.
                & 120 \\
            \textit{Lengths}
                & Single, large length constraint on a variable.
                & 200 \\
            \textit{Lengths-Concats}
                & Tree of fixed-length concats of variables.
                & 100 \\
            \textit{Overlaps}
                & Formula of the form A.X = X.B.
                & 80 \\
            \textit{Regex}
                & Complex regex membership test.
                & 120 \\
            \textit{Many-Regexes}
                & Multiple random regex membership tests.
                & 40 \\
            \textit{Regex-Deep}
                & Regex membership test with many nested operators.
                & 45 \\
            \textit{Regex-Pair}
                & Test for membership in one regex, but not another.
                & 40 \\
            \textit{Regex-Lengths}
                & Regex membership test, and a length constraint.
                & 40 \\
            \textit{Different-Prefix}
                & Equality of two deep concats with different prefixes.
                & 60 \\
            \bottomrule
        \end{tabular}
    \end{subtable}
